---
title: "Astronauts"
output: html_notebook
---

```{r astronaut-import-libraries, include = F}
x <- c("igraph", "dplyr", "magrittr")
sapply(x, library, character.only = T)
```

This project was based on the astronaut [dataset](https://www.kaggle.com/jessemostipak/astronaut-database) at kaggle. It was downloaded and imported. There were `r nrow(df)` rows and `r ncol(df)` columns.

```{r astronaut-import-data}
input <- "./data/astronauts.csv"
df <- data.table::fread(input = input)
```

# Preliminaries

## Missingness

Let's check for missingness. `r length(which(is.na(df) == T))` values are missing.  Subsetting to only the rows that contain a missing value yields the dataframe below.

```{r astro-subset-nas}
df.na <- df[!complete.cases(df), ]
df.na
```

Also, let's understand what values in the variables are missing and see if there is any discernible pattern.

```{r astro-amelia-missmap}
Amelia::missmap(df.na)
```

## Unique values

Also, let's count the unique values within each column to identify potential factor variables. The `id` field is clearly a unique identifier for each row. It also appears that each astronaut has a unique identifier in the `number` field in addition to the `name` field.

```{r astro-unique-values}
rev(sort(apply(df, 2, function(x) length(unique(x)))))
```

## Variable Types

```{r astro-column-classes}
class(df$id)
table(unlist(lapply(df, class)))
```

## Summary

```{r}
my.summary <- do.call(cbind, lapply(df, summary, digits = 2))
my.summary
```

## Challenges

There was some difficulty in the data which was discovered when creating a network analysis.  The names were problematic. "Aleksandrov, Aleksandr", a Russian astronaut, is the same person but appears in the astronaut subset twice.  The other issue is that two astronauts have been assigned the number "385": Yuri Baturin and Pedro Duque.  This challenge will be set aside for the time being.

```{r astronaut-subset-name}
df2 <- df[, c(2, 4)] #subset main df to just names and number
df2 <- dplyr::distinct(df2) #filter to unique/distinct values 564
length(unique(df2$name)) #564 astronauts
length(unique(df2$number))#564 unique numbers
#one astronaut is included in the names column twice--it is the same person
duplicate.name <- df2$name[duplicated(df2$name)] #"Aleksandrov, Aleksandr"
df2[grep("Aleksandrov, Aleksandr", df2$name)]
#two astronauts have the same number
duplicate.number <- df2$number[duplicated(df2$number)]
df[which(df$number == duplicate.number)]
```

```{r }
df3 <- dplyr::select_if(df, is.numeric)
df3 <- dplyr::select(df3, -id, -number, -mission_number)
res <- cor(df3)
res <- round(res, 2)
library(corrplot)
col<- colorRampPalette(c("blue", "white", "red"))(20)
par(mar = c(3, 3, 3, 3))
heatmap(x = res, col = col, 
        symm = TRUE, 
        margins = c(5, 5),
        main = "Numeric Astonaut Variables")
```

